Location and Scale-Invariant Power Transformations for Transforming Data to Normality

### 2024-08-29 manuscript/manuscript.Rmd
Replaced k_1 and k_2 by delta_1 and delta_2.

- A step function, with $k_1 \geq 0$ as threshold parameter:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| \leq k_1\\
0 & \text{if } \left| \dot{x}_i \right| > k_1
\end{cases}
\end{equation}
```

- A triangle function (or generalised Huber weight), with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < k_1\\
1 - \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
0 & \text{if } \left| \dot{x}_i \right| > k_2
\end{cases}
\end{equation}
```

- A tapered cosine function [@Tukey1967-eb], with $k_1 \geq 0$ and $k_2 \geq k_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < k_1\\
0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - k_1}{k_2 - k_1} \right) & \text{if } k_1 \leq \left| \dot{x}_i \right| \leq k_2 \\
0 & \text{if } \left| \dot{x}_i \right| > k_2
\end{cases}
\end{equation}
```

All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< k_1$, instances are fully weighted, i.e. when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top.
The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.
Examples of these weighting function are shown in Figure \ref{fig:weighting-functions}.

```{r weighting-functions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=2.5}
cap <- paste0(
  "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
  "The step function was parameterised with $k_1 = 0.60$. ",
  "The triangle and tapered cosine functions were both parameterised with $k_1 = 0.30$ and $k_2 = 0.90$.")

@@

To find the optimal values for the weighting function parameters $k_1$ and $k_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:

```{=latex}
\begin{equation}
\label{eqn:minimisation-weighting}
\left\{ \hat{k}_1, \hat{k}_2 \right\} = \argmin_{k_1, k_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; k_1, k_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
\end{equation}
```

@@

\toprule
method & \multicolumn{3}{c}{$k_1$} & \multicolumn{3}{c}{$k_2$} & loss \\
& initial & limits & optimal & initial & limits & optimal & \\

@@

# Code availability

Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository. The manuscript was created using R Markdown and is also available from the power.transform GitHub repository.

>>>>

- A step function, with $\delta_1 \geq 0$ as threshold parameter:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| \leq \delta_1\\
0 & \text{if } \left| \dot{x}_i \right| > \delta_1
\end{cases}
\end{equation}
```

- A triangle function (or generalised Huber weight), with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
1 - \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}
```

- A tapered cosine function [@Tukey1967-eb], with $\delta_1 \geq 0$ and $\delta_2 \geq \delta_1$ as threshold parameters:
```{=latex}
\begin{equation}
w_i =
\begin{cases}
1 & \text{if } \left| \dot{x}_i \right| < \delta_1\\
0.5 + 0.5 \cos\left(\pi \frac{\left| \dot{x}_i \right| - \delta_1}{\delta_2 - \delta_1} \right) & \text{if } \delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2 \\
0 & \text{if } \left| \dot{x}_i \right| > \delta_2
\end{cases}
\end{equation}
```

All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< \delta_1$, instances are fully weighted, i.e. when $\delta_1 > 0$ the weighting functions are symmetric window functions with a flat top.
The triangle and tapered cosine functions then gradually down-weight instances with $\delta_1 \leq \left| \dot{x}_i \right| \leq \delta_2$, and assign no weight to instances $\left| x_i \right| > \delta_2$.
Examples of these weighting function are shown in Figure \ref{fig:weighting-functions}.

```{r weighting-functions, warning=FALSE, message=FALSE, fig.cap=cap, fig.height=1.5}
cap <- paste0(
  "Weighting functions investigated in this study to make power transformations more robust against outliers. ",
  "The step function was parameterised with $\\delta_1 = 0.60$. ",
  "The triangle and tapered cosine functions were both parameterised with $\\delta_1 = 0.30$ and $\\delta_2 = 0.90$.")

@@

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable), we minimised the absolute difference between the $\lambda_{r}$ parameter obtained for transformation in the presence of outliers, and the $\lambda_0$ parameter obtained using the non-robust transformations in absence of outliers:

```{=latex}
\begin{equation}
\label{eqn:minimisation-weighting}
\left\{ \hat{\delta}_1, \hat{\delta}_2 \right\} = \argmin_{\delta_1, \delta_2} \sum_{i=1}^{m_d} \sum_{j=1}^{m_{out}} \left| \lambda_r \left(\mathbf{X}_{i, j}; \delta_1, \delta_2 \right) - \lambda_0 \left(\mathbf{X}_i \right) \right|
\end{equation}
```

@@

\toprule
method & \multicolumn{3}{c}{$\delta_1$} & \multicolumn{3}{c}{$\delta_2$} & loss \\
& initial & limits & optimal & initial & limits & optimal & \\

@@

# Data and code availability

Location- and scale-invariant power transformations are implemented in the `power.transform` package for R, and is available from [https://github.com/oncoray/power.transform] and the CRAN repository [INSERT LINK HERE]. The manuscript was created using R Markdown and is also available from the power.transform GitHub repository. Data and results for the machine learning experiment are separately available from Zenodo [INSERT LINK HERE].
